\section*{Acknowledgements}
We would like to express our sincere gratitude to the following individuals and organizations who have contributed to the successful completion of this thesis:
\begin{itemize}
    \item Pierre Schaus, our promotor, for his valuable guidance, supervision, and continuous support throughout the research process. His insights, enthusiasm, constructive advice, and feedback have been decisive in shaping this work.
    \item Achille Morenville for his technical guidance and his follow-up throughout the whole year. He has been available at every stage of this master's thesis. We are very thankful for his advice and willingness to share his insights and knowledge.
    \item Timothée Lonfils from Buildwise, for his trust and willingness to help us and valuable insights into the market and industry practices. His expertise and input have provided valuable context and relevance to the research.
    \item Buildwise, for providing access to the robot SPOT. Their support and collaboration have allowed us to collect data crucial to this study.
\end{itemize}

Also, we would like to extend our gratitude to all the professors of UCLouvain for their dedication and their contributions to our education. Their passion for knowledge and commitment to excellence have played an essential role in shaping our understanding of the subject matter.\\

Additionally, we would like to acknowledge our friends and family for their support, understanding, and encouragement throughout this journey.\\

Lastly, we would like to express our appreciation to everyone else who has supported us directly or indirectly during this thesis. Your support, whether big or small, has made a significant difference in the completion of this work.